---
title: "covid 19"
author: "Kimberly Lemus"
date: "summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```
## Graphic

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
```
In this boxplot, we can see that more women are alive than men. Women are more prominent than men. The amount of women and men that died does not substantial. All and all, there were more men that died compared to women. 

  
```{r}
barchartGC(~result + Sex,data=kRdata, type="percent")
```

In the second plot, it shows that the percentage of females who are alive is more than the percentage of males that are alive. As for the dead, it shows that the percentage of males that are dead is about twice as higher as the percentage of females that are dead.


  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
 

## Inferential results 









## Conclusions






