---
title: "covid 19"
author: "Kimberly Lemus"
date: "summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```
## Graphic

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
```
In this boxplot, we can see that more women are alive than men. Women are more prominent than men. The amount of women and men that died does not substantial. All and all, there were more men that died compared to women. 

  
```{r}
barchartGC(~result + Sex,data=kRdata, type="percent")
```

In the second plot, it shows that the percentage of females who are alive is more than the percentage of males that are alive. As for the dead, it shows that the percentage of males that are dead is about twice as higher as the percentage of females that are dead.


  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
 
 We now have the correct percentages, where we can see the percentage of the alive are Female 56% and Male 44%. The percentage of dead are 64% for Male and 36% for Female. We can see there is a backwards relationship between males and females. Females are the predominantly alive group where males were predominantly the dead group. Of all the females 1.4% ended up dead, compared to males, where 3% ended up dead. Males deaths were doubled compared to the rate of women. 
 
 
 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
```

The chi sq test provides the exact numbers. It shows that 1831 were females and 1411 were males are in the alive group. 25 females died vs 44 males died. The odds of a female dying is (25/1831) and the odds of a male dying is (44/1411). The odds ratio of a male dying to a female is (44/1411)/(25/1831). The odds ratio is 2.3.


```{r}
fisher.test(table1)
```

The fisher test gives us the odd ratio, which we also got by doing the chi sq test. We also see the confidence intervals are 1.4 to 3.9, meaning the odd ratio can go as low as 1.4 or as high as 3.9. 

## Conclusions

In conclusion, we can see a big statistical difference between males and females vs COVID. Based off our results, more males die of COVID vs females. Our p-value is very small, so therefore we reject the null hypothesis. More importantly we can see that more males die from this disease vs females.




